---
title: "p8105_hw2_10032019"
author: "Kodiak Soled"
date: "10/3/2019"
output: github_document
always_allow_html: yes
---

_Note to TAs: you may need to_ `install.packages(kableExtra)` _to run my code for problem #3._

# *Problem #1*

## _Mr. Trash Wheel_

### First, we can load the `readxl` package to import the Excel file data and we can load the `tidyverse` package for some of the datacleaning we will need to perform: 
```{r}
library(readxl)
library(tidyverse)
```

### Now we can read (`read_excel`) and clean the dataset by: 
* Specifying the sheet in the Excel file we want with `sheets`
* Omitting non-data entries by specifying the columns and rows we want to include with the `range` function
* Cleaning up the names of the variables with the `janitor::clean_names()` function
* Renaming the variables (`rename`) to have simplier names
* Omitting the rows that do not include dumpster-specific data by using `drop_na`

### Finally we can round the number of sports balls to the nearest integer and convert the result to an integer variable using `as.integer`. 
```{r}
mr_trash_wheel = read_excel("./data/TrashWheelCollectionTotals.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N406") %>%
  janitor::clean_names() %>%
  rename(weight = weight_tons, volume = volume_cubic_yards) %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(sports_balls, 0))
    
mr_trash_wheel
```

## _Precipitation Data_

### First, we can read (`read_excel`) and clean (`janitor::clean_names()`) the preciptation data for the two datasheets that contains 2017 and 2018 data. Again, we can omit the rows without the precipitation data by specifying the `range` of rows and columns to include. Lastly, we can add a variable year to each data set with the `mutate` function. 
```{r}
precip_data_2018 = 
  read_excel("./data/TrashWheelCollectionTotals.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2018)

precip_data_2018
  
precip_data_2017 = 
  read_excel("./data/TrashWheelCollectionTotals.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2017)
 
precip_data_2017
``` 

### Finally, we can combine the two datasets with the `bind_rows` function to stack the two datasets ontop of one another to create a tidy dataset. We can also convert the month variable from a numeric (1-12) to a character variable (january-december) so that the month names appear using `month.name`:
```{r message = FALSE}
precip_data = 
  bind_rows(precip_data_2017, precip_data_2018) %>%
  mutate(
    month = month.name[month],
    month = str_to_lower(month))

precip_data
```

## _Description_
### Mr. Trash Wheel 

* The Mr. Trash Wheel dataset contains data from May 2014 until June 2019. 
* The dataset includes `r nrow(mr_trash_wheel)` observations and `r ncol(mr_trash_wheel)` variables for a total of `r nrow(mr_trash_wheel)*ncol(mr_trash_wheel)` cells. 
* The `r length(mr_trash_wheel)` variables include: dumpster, month, year, date, weight, volume, etc. 
* The median number of sports balls in a dumpster in 2017 was `r median(pull(mr_trash_wheel %>% filter(year == "2017"), sports_balls))`.

### Precipitation

* The combined precipition dataset contains the total preciptation per month for the years 2017 and 2018. 
* It contains `r nrow(precip_data)` observations and `r ncol(precip_data)` variables (month, year, and total precipitation) for a total of `r nrow(precip_data)*ncol(precip_data)` cells.
* In 2018 the total precipitation was `r sum(pull(precip_data %>% filter(year == "2018"), total))`. 

# *Problem 2* 

## _FiveThiryEight_

### First, we can import (`read_csv`) and clean (`janitor::clean_names()`) the three datasets: pols-month.csv, unemployment.csv, and snp.csv. Then, for the pols-month and snp datasets, we can use the `separate` function to change the "date" variable into three variables ("year", "month", and "day") (_note: this isn't necessary for the unemployment dataset as it was already seperated by date_). Then we can make specific modifications to each dataset in order to merge them: 

* For the pols-month data, we can replace the month number with the abbreviated month name (to match the unemployment dataset that contains abbreviated month names) with `month.abb` under the `mutate` function, after specifying the month as an integer (`as.integer`). We also can create a new variable "president" under `mutate` with the `if_else` function and remove the "prez_gap", "prez_day", and "day" variables with `select` to clean up the dataset. 

* We can then clean the snp data to look similar to the pols-month data by using many of the same functions as above (e.g., deleting the "day" variable with `select`, changing month as a number to a name with `month.abb` under the `mutate` function, etc.)  

* The unemployment data needed to be reorganized from a "wide" to "long" format to match the first two datasets. The `pivot_longer` function allows us to do this. We also needed to make"year" into a character vector which we can do with `as.character` under the `mutate` function. 

### The cleaned datasets look like the following: 
```{r message = FALSE}
pols_month =
  read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(
    month = month.abb[as.integer(month)],
    month = str_to_lower(month),
    president = if_else(prez_gop == 1, "gop", "dem")) %>%
  select(-prez_gop, -prez_dem, -day)

pols_month

snp = 
  read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year")) %>%
  select(-day) %>%
  mutate(
    month = month.abb[as.integer(month)],
    month = str_to_lower(month)) %>%
  select(year, month, close) %>%
  arrange(year, month) %>%
  rename(snp_close = close)

snp

unemployment = 
  read_csv("./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "count") %>%
  mutate(year = as.character(year)) %>%
  rename(unemployment_rate = count)

unemployment
```

### Now we are ready to combine the three datasets. We can first merge the pols-month and snp datasets with the `left_join` function and specifying the merge by the shared catagories of "month" and "year". Then, we can merge this new dataset with the third dataset unemployment in the same fashion: 
```{r}
pol_snp_data = 
  left_join(pols_month, snp, by = c("month", "year"))

pol_snp_unemployment_data = 
  left_join(pol_snp_data, unemployment, by = c("month", "year")) %>%
  select(month, year, president, unemployment_rate, snp_close, everything())

pol_snp_unemployment_data
```