---
title: "p8105_hw2_10032019"
author: "Kodiak Soled"
date: "10/3/2019"
output: github_document
always_allow_html: yes
---

_Note to TAs: you may need to_ `install.packages(kableExtra)` _to run my code for problem #3._

# *Problem #1*

## _Mr. Trash Wheel_

### First, we can load the `readxl` package to import the Excel file data and we can load the `tidyverse` package for some of the datacleaning we will need to perform: 
```{r}
library(readxl)
library(tidyverse)
```

### Now we can read (`read_excel`) and clean the dataset by: 
* Specifying the sheet in the Excel file we want with `sheets`
* Omitting non-data entries by specifying the columns and rows we want to include with the `range` function
* Cleaning up the names of the variables with the `janitor::clean_names()` function
* Renaming the variables (`rename`) to have simplier names
* Omitting the rows that do not include dumpster-specific data by using `drop_na`

### Finally we can round the number of sports balls to the nearest integer and convert the result to an integer variable using `as.integer`. 
```{r}
mr_trash_wheel = read_excel("./data/TrashWheelCollectionTotals.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N406") %>%
  janitor::clean_names() %>%
  rename(weight = weight_tons, volume = volume_cubic_yards) %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(sports_balls, 0))
    
mr_trash_wheel
```

## _Precipitation Data_

### First, we can read (`read_excel`) and clean (`janitor::clean_names()`) the preciptation data for the two datasheets that contains 2017 and 2018 data. Again, we can omit the rows without the precipitation data by specifying the `range` of rows and columns to include. Lastly, we can add a variable year to each data set with the `mutate` function. 
```{r}
precip_data_2018 = 
  read_excel("./data/TrashWheelCollectionTotals.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2018)

precip_data_2018
  
precip_data_2017 = 
  read_excel("./data/TrashWheelCollectionTotals.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2017)
 
precip_data_2017
``` 